% napkin/biblio.tex created 20190316
% The following statement selects the style to use for references.  It controls the sort order of the entries in the bibliography and also the formatting for the in-text labels.
\bibliographystyle{plain}
% This specifies the location of the file containing the bibliographic information.  
% It assumes you're using BibTeX (if not, why not?).
%\cleardoublepage % This is needed if the book class is used, to place the anchor in the correct page,
                 % because the bibliography will start on its own page.
                 % Use \clearpage instead if the document class uses the "oneside" argument
%\phantomsection  % With hyperref package, enables hyperlinking from the table of contents to bibliography             
% The following statement causes the title "References" to be used for the bibliography section:
%\renewcommand*{\bibname}{References}

% Add the References to the Table of Contents
\addcontentsline{toc}{section}{References}

\begin{thebibliography}{}

\bibitem{And}
Andrews and Eriksson, \emph{Integer Partitions}.

\bibitem{And2}
Andrews, \emph{Theory of Partitions}.

\bibitem{And3}
Andrews, `Applications of Basic Hypergeometric Functions'.
% http://www.personal.psu.edu/gea1/pdf/59.pdf

\bibitem{Ati} % 20190317
M.\ Atiyah, `The Logarithm of the Dedekind $\eta$-Function'. 1987.

\bibitem{Baez183} % 20190318
John Baez, \emph{This Week's Finds}, \texttt{week183} onward.

\bibitem{Ber} % 201903
Bergeron, Labelle, Leroux, \emph{Introduction to the Theory of Species of Structures}.

\bibitem{Ber2} % 20190321
Bergeron, Labelle, Leroux, \emph{Combinatorial Species and Tree-Like Structures}. 
% main ref. for theory of species

\bibitem{Chen} % 20190318
E.\ Chen, \emph{Napkin}. \url{www.evanchen.cc}.

\bibitem{Des}
W.\ Deskins, \emph{Abstract Algebra}.

\bibitem{Efr} % 20190318
Efron and Hastie, \emph{Computer Age Statistical Inference}.

\bibitem{Engl} % 20190318
Engler and Prestel, \emph{Valued Fields}.

\bibitem{Foll} % 20190318
Folland, \emph{Quantum Field Theory}.

\bibitem{Froh} % 20190318
Fr\"ohlich.

\bibitem{Gan} % 20190321
T.\ Gannon, \emph{Moonshine Beyond the Monster}.

\bibitem{God1} % 20190317
C.\ Godsil, \emph{Introduction to Combinatorics} (\texttt{enum2.pdf}). Course notes used for MATH 249.

\bibitem{Hat} % 20190318
A.\ Hatcher, \emph{Algebraic Topology}.

\bibitem{Harr1} % 20190318
Joe Harris, \emph{Introduction to the Moduli Space of Curves}?

\bibitem{Jay} % 20190318
Jaynes, first paper on statistical mechanics and information theory.

\bibitem{Jon} % 20190320
Jones and Singerman, \emph{Complex Functions: An Algebraic and Geometric Viewpoint}. 1987.

\bibitem{KLM} % 20190318
Kaye, Laflamme, and Mosca. Quantum computation book.

\bibitem{Kac}
Kac and Cheung, \emph{Quantum Calculus}.
% Faulhaber's formula and Bernoulli numbers/polys, partition identities, Jacobi triple product, q-binomials, q-Hypergeometric functions...

\bibitem{Knu} % 20190318
D.\ Knuth, \emph{Concrete Mathematics}.

\bibitem{Land} % 20190318
Lando and Zvonkin, \emph{Graphs on Surfaces and their Applications}.

\bibitem{Lax} % 20190318
P.\ Lax, \emph{Linear Algebra and its Applications}.

\bibitem{Mac1} % 20190318
\emph{Sheaves in Geometry and Logic}.

\bibitem{May} % 20190318
May, \emph{A Concise course on algebraic Topology}.

\bibitem{nLab} % 20190318
nLab.

\bibitem{nLab2} % 20190321
\url{https://ncatlab.org/nlab/show/too+simple+to+be+simple}.

\bibitem{Neu} % 20190318
Neukirch, \emph{Algebraic Number Theory}.

\bibitem{Niel} % 20190318
Nielsen and Chuang. Quantum information book.

\bibitem{PM351}
PMATH 351 course notes.

\bibitem{PM440} % 20190318
PMATH 440 course notes. \url{https://uwaterloo.ca/pure-mathematics/sites/ca.pure-mathematics/files/uploads/files/pmath440notes_0.pdf}.

\bibitem{ssp} % 20190318
\texttt{ssp} blog.

\bibitem{S14sem} % 20190318
\textsf{mlbaker}, Summer 2014 seminar on semisimple Lie theory etc.

\bibitem{Sard} % 20190318
G.\ Sardanashvily.

\bibitem{Sarn} % 20190318
P.\ Sarnak, `Some Applications of Modular Forms'.

\bibitem{Sch} % 20190318
Scholkopf and Smola. Book on kernel methods.

\bibitem{Shar} % 20190318
Romyar Sharifi, ``Iwasawa Theory: A Climb Up The Tower'' in the AMS Notices (January 2019).

\bibitem{Span} % 20190318
E.\ Spanier, \emph{Algebraic Topology}.

\bibitem{Sta1}
R.\ P.\ Stanley, \emph{Enumerative Combinatorics} Volume 1.

\bibitem{Sta2}
R.\ P.\ Stanley, \emph{Enumerative Combinatorics} Volume 2.

\bibitem{Ter} % 20190317
A.\ Terras, \emph{Harmonic Analysis on Symmetric Spaces: Euclidean Space, the Sphere, and the Poincar\'e Upper Half-Plane}.

\bibitem{Tib1} % 20190318
Tibshirani et al, \emph{Elements of Statistical Learning}.

\bibitem{Tib2} % 20190318
Tibshirani et al, \emph{Introduction to Statistical Learning}.

\bibitem{Vile} % 20190318
Vilenkin and Klimyk, \emph{Representation of Lie Groups and Special Functions} (1991--1993). 3 volumes.

\bibitem{Watr} % 20190318
John Watrous. New book on quantum information theory.

\bibitem{Whit} % 20190318
Whittaker and Watson, \emph{A Course in Modern Analysis}.

\bibitem{Yos} % 20190317
Masaaki Yoshida, \emph{Hypergeometric Functions, My Love}.

\end{thebibliography}